### 2020-06-27 latex/chapter-skippable.tex
new chapter-skippable.tex

\chapter[Skippable chapter]{A skippable introduction to ordinal numbers and the Coq proof assistant}

blah.    blah.

\section{blah}