### 2021-09-20 doc/JFLA2022/Ex42b.tex
small corrections

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      ~~\PY{n+nb}{assert}~\PY{o}{(}\PY{n}{HAP}\PY{o}{:}~\PY{n}{In}~\PY{n}{AP}~\PY{o}{(}\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{))}~\PY{n+nb+bp}{by}~\PY{n+nb}{apply}~\PY{n}{AP\PYZus{}phi0}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~\PY{n+nb}{elim}~~\PY{n}{HAP}\PY{o}{;}~\PY{n+nb}{intros}~~\PY{n}{\PYZus{}}~\PY{n}{H0}\PY{o}{;}~\PY{n+nb}{apply}~\PY{n}{H0}\PY{o}{;}~\PY{n+nb}{clear}~\PY{n}{H0}\PY{o}{.}
    \end{input}
    \sep
    \begin{output}
      \begin{goals}
        \begin{goal}
          \begin{hyps}
            \hyp{HAP}{\PY{n}{In}~\PY{n}{AP}~\PY{o}{(}\PY{n}{phi0}~\PY{o}{(}\PY{n}{F}~\PY{l+m+mi}{2}\PY{o}{))}}
          \end{hyps}
          \sep
          \infrule{}
          \sep
          \begin{conclusion}
            \PY{n+nb+bp}{omega}~\PY{o}{+}~\PY{n}{F}~\PY{l+m+mi}{42}~\PY{o}{\PYZlt{}}~\PY{n}{phi0}~\PY{o}{(}\PY{n}{F}~\PY{l+m+mi}{2}\PY{o}{)}
          \end{conclusion}
        \end{goal}
      \end{goals}
    \end{output}
  \end{sentence}
\end{alectryon}