### 2021-10-04 doc/JFLA2022/snippets/E0/Ex42.tex

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Example}~\PY{n+nf}{Ex42}\PY{o}{:}~\PY{o}{(}\PY{n}{omega}~\PY{o}{+}~\PY{l+m+mi}{42}~\PY{o}{+}~\PY{n}{omega}\PY{o}{\PYZca{}}~\PY{l+m+mi}{2}~\PY{o}{=}~\PY{n}{omega}\PY{o}{\PYZca{}}~\PY{l+m+mi}{2}\PY{o}{)\PYZpc{}}\PY{n}{e0}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Proof}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~\PY{n+nb}{rewrite}~\PY{o}{\PYZlt{}\PYZhy{}}~~\PY{n}{Comparable}\PY{o}{.}\PY{n}{compare\PYZus{}eq\PYZus{}iff}\PY{o}{.}
    \end{input}
    \sep
    \begin{output}
      \begin{goals}
        \begin{goal}
          \begin{hyps}\end{hyps}
          \sep
          \infrule{}
          \sep
          \begin{conclusion}
            \PY{n+nb}{compare}~\PY{o}{(}\PY{n}{omega}~\PY{o}{+}~\PY{l+m+mi}{42}~\PY{o}{+}~\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{)}~\PY{o}{(}\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{)}~\PY{o}{=}~\PY{n}{Eq}
          \end{conclusion}
        \end{goal}
      \end{goals}
    \end{output}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~\PY{n+nb+bp}{reflexivity}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Qed}\PY{o}{.}
    \end{input}
  \end{sentence}
\end{alectryon}