### 2021-10-05 doc/JFLA2022/snippets/Hprime/HprimeDef.tex

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      \PY{n}{Equations}~\PY{n}{H\PYZsq{}\PYZus{}}~\PY{o}{(}\PY{n}{alpha}\PY{o}{:}~\PY{n}{E0}\PY{o}{)}~\PY{o}{(}\PY{n}{i}\PY{o}{:}\PY{n}{nat}\PY{o}{)}~\PY{o}{:}~~\PY{n}{nat}~~\PY{n+nb+bp}{by}~\PY{k+kn}{wf}~~\PY{n}{alpha}~\PY{n}{Lt}~\PY{o}{:=}\nl
      ~~\PY{n}{H\PYZsq{}\PYZus{}}~\PY{n}{alpha}~~\PY{n}{i}~\PY{k+kr}{with}~\PY{n}{E0\PYZus{}eq\PYZus{}dec}~\PY{n}{alpha}~\PY{n}{Zero}~\PY{o}{:=}\nl
      ~~~~\PY{o}{\PYZob{}}~\PY{o}{|}~\PY{n+nb}{left}~\PY{n}{\PYZus{}zero}~\PY{o}{=\PYZgt{}}~~\PY{n}{i}~\PY{o}{;}\nl
      ~~~~~~\PY{o}{|}~\PY{n+nb}{right}~\PY{n}{\PYZus{}nonzero}\nl
      ~~~~~~~~~~\PY{k+kr}{with}~\PY{n}{Utils}\PY{o}{.}\PY{n}{dec}~\PY{o}{(}\PY{n}{Limitb}~\PY{n}{alpha}\PY{o}{)}~\PY{o}{:=}\nl
      ~~~~~~~~~~\PY{o}{\PYZob{}}~\PY{o}{|}~\PY{n+nb}{left}~\PY{n}{\PYZus{}limit}~\PY{o}{=\PYZgt{}}~~\PY{n}{H\PYZsq{}\PYZus{}}~\PY{o}{(}\PY{n}{Canon}~\PY{n}{alpha}~\PY{o}{(}\PY{n}{S}~\PY{n}{i}\PY{o}{))}~~\PY{n}{i}~\PY{o}{;}\nl
      ~~~~~~~~~~~~\PY{o}{|}~\PY{n+nb}{right}~\PY{n}{\PYZus{}successor}~\PY{o}{=\PYZgt{}}~~\PY{n}{H\PYZsq{}\PYZus{}}~\PY{o}{(}\PY{n}{Pred}~\PY{n}{alpha}\PY{o}{)}~\PY{o}{(}\PY{n}{S}~\PY{n}{i}\PY{o}{)\PYZcb{}\PYZcb{}.}
    \end{input}
  \end{sentence}
\end{alectryon}