### 2022-01-05 doc/JFLA2022/Slides/ONDef.tex

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Class}~\PY{n+nf}{ON}~\PY{o}{\PYZob{}}\PY{n+nv}{A}\PY{o}{:}\PY{k+kt}{Type}\PY{o}{\PYZcb{}}~\PY{o}{(}\PY{n+nv}{lt}\PY{o}{:}~\PY{n}{relation}~\PY{n}{A}\PY{o}{)}~\PY{o}{(}\PY{n+nv}{cmp}\PY{o}{:}~\PY{n}{Compare}~\PY{n}{A}\PY{o}{)}~\PY{o}{:=}\nl
      ~~\PY{o}{\PYZob{}}\nl
      ~~\PY{n}{ON\PYZus{}comp}~\PY{o}{:\PYZgt{}}~\PY{n}{Comparable}~\PY{n}{lt}~\PY{n}{cmp}\PY{o}{;}\nl
      ~~\PY{n}{ON\PYZus{}wf}~\PY{o}{:}~\PY{n}{well\PYZus{}founded}~\PY{n}{lt}\PY{o}{;}\nl
      ~~\PY{o}{\PYZcb{}.}
    \end{input}
  \end{sentence}
\end{alectryon}