### 2022-01-29 doc/JFLA2022/Slides/multA.tex
update slides about Gaia

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Lemma}~\PY{n+nf}{mult\PYZus{}refR}~\PY{o}{(}\PY{n+nv}{a}~\PY{n+nv}{b}~\PY{o}{:}~\PY{n}{T1}\PY{o}{)}~\PY{o}{:}~\PY{n}{hmult}~\PY{o}{(}\PY{n}{g2h}~\PY{n}{a}\PY{o}{)}~\PY{o}{(}\PY{n}{g2h}~\PY{n}{b}\PY{o}{)}~\PY{o}{=}~\PY{n}{g2h}~\PY{o}{(}\PY{n}{a}~\PY{o}{*}~\PY{n}{b}\PY{o}{).}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Proof}\PY{o}{.}~
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{n+nb}{apply}~\PY{n}{refines2\PYZus{}R}\PY{o}{,}~~\PY{n}{mult\PYZus{}ref}\PY{o}{.}~
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Qed}\PY{o}{.}~\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{txt}
    \nl
  \end{txt}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Lemma}~\PY{n+nf}{hmultA}~\PY{o}{:}~\PY{n}{associative}~\PY{n}{hmult}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Proof}\PY{o}{.}~\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~\PY{n+nb}{move}~\PY{o}{=\PYZgt{}}~\PY{n}{a}~\PY{n}{b}~\PY{n}{c}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~\PY{n+nb+bp}{by}~\PY{n+nb}{rewrite}~\PY{o}{\PYZhy{}(}\PY{n}{g2h\PYZus{}h2g}~\PY{n}{a}\PY{o}{)}~\PY{o}{\PYZhy{}(}\PY{n}{g2h\PYZus{}h2g}~\PY{n}{b}\PY{o}{)}~\PY{o}{\PYZhy{}(}\PY{n}{g2h\PYZus{}h2g}~\PY{n}{c}\PY{o}{)}~\PY{o}{!}\PY{n}{mult\PYZus{}refR}~\PY{n}{mulA}\PY{o}{.}~\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Qed}\PY{o}{.}
    \end{input}
  \end{sentence}
\end{alectryon}