### 2022-01-30 doc/JFLA2022/Slides/canonDef.tex
Update slides (#118)
* update slides about Gaia

* new snippets for gaia-hydras

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Definition}~\PY{n+nf}{canon}~\PY{o}{(}\PY{n+nv}{a}\PY{o}{:}~\PY{n}{T1}\PY{o}{)}~\PY{o}{(}\PY{n+nv}{i}\PY{o}{:}\PY{n}{nat}\PY{o}{)}~\PY{o}{:}~\PY{n}{T1}~\PY{o}{:=}\nl
      ~~\PY{n}{h2g}~\PY{o}{(}\PY{n}{hcanon}~\PY{o}{(}\PY{n}{g2h}~\PY{n}{a}\PY{o}{)}~\PY{n}{i}\PY{o}{).}
    \end{input}
  \end{sentence}
\end{alectryon}