### 2022-01-30 doc/JFLA2022/Slides/gcanonLimitOf.tex

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      ~\PY{k+kn}{Lemma}~\PY{n+nf}{gcanon\PYZus{}limit\PYZus{}of}~\PY{n+nv}{lambda}~\PY{o}{(}\PY{n+nv}{Hnf}~\PY{o}{:}~\PY{n}{T1nf}~\PY{n}{lambda}\PY{o}{)}~\PY{o}{(}\PY{n+nv}{Hlim}~\PY{o}{:}~\PY{n}{T1limit}~\PY{n}{lambda}\PY{o}{)}~\PY{o}{:}\nl
      ~~~\PY{n}{limit\PYZus{}of}~\PY{o}{(}\PY{n}{canon}~\PY{n}{lambda}\PY{o}{)}~\PY{n}{lambda}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~\PY{k+kn}{Proof}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~~\PY{n+nb}{split}~\PY{o}{=\PYZgt{}}~\PY{o}{//.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~~\PY{n+nb}{move}~\PY{o}{=\PYZgt{}}~\PY{n}{n}~\PY{n}{m}~\PY{n}{Hnm}\PY{o}{;}~\PY{n+nb}{apply}~\PY{n}{gcanon\PYZus{}limit\PYZus{}mono}~\PY{o}{=\PYZgt{}}~\PY{o}{//.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~~\PY{n+nb}{apply}~\PY{n}{gcanon\PYZus{}limit\PYZus{}v2}~\PY{o}{=\PYZgt{}}~\PY{o}{//.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~\PY{k+kn}{Qed}\PY{o}{.}~~~~
    \end{input}
  \end{sentence}
\end{alectryon}