### 2022-06-02 doc/JFLA2022/Slides/gcanonSLE.tex

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Lemma}~\PY{n+nf}{canonS\PYZus{}LE}~\PY{n+nv}{a}~\PY{n+nv}{n}\PY{o}{:}\nl
      ~~~~\PY{n}{T1nf}~\PY{n}{a}~\PY{o}{\PYZhy{}\PYZgt{}}~\PY{n}{canon}~\PY{n}{a}~\PY{n}{n}\PY{o}{.+}\PY{l+m+mi}{1}~\PY{o}{\PYZlt{}=}~\PY{n}{canon}~\PY{n}{a}~\PY{n}{n}\PY{o}{.+}\PY{l+m+mi}{2}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Proof}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~\PY{n+nb}{rewrite}~\PY{o}{\PYZhy{}(}\PY{n}{h2g\PYZus{}g2hK}~\PY{n}{a}\PY{o}{)}~\PY{o}{\PYZhy{}}\PY{n}{nf\PYZus{}ref}~\PY{o}{/}\PY{n}{canon}~\PY{n}{g2h\PYZus{}h2gK}~\PY{o}{=\PYZgt{}}~\PY{n}{Hnf}\PY{o}{;}\nl
      ~~\PY{n+nb+bp}{by}~\PY{n+nb}{apply}~\PY{n}{le\PYZus{}ref}\PY{o}{,}~\PY{n}{canonS\PYZus{}LE}\PY{o}{.}~\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Qed}\PY{o}{.}
    \end{input}
  \end{sentence}
\end{alectryon}