### 2022-06-02 doc/JFLA2022/Slides/gcanonLimitOf.tex
small changes to JFLA slides

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      ~\PY{k+kn}{Lemma}~\PY{n+nf}{canon\PYZus{}limit\PYZus{}of}~\PY{n+nv}{lambda}~\PY{o}{(}\PY{n+nv}{Hnf}~\PY{o}{:}~\PY{n}{T1nf}~\PY{n}{lambda}\PY{o}{)}~\PY{o}{(}\PY{n+nv}{Hlim}~\PY{o}{:}~\PY{n}{T1limit}~\PY{n}{lambda}\PY{o}{)}~\PY{o}{:}\nl
      ~~~\PY{n}{limit\PYZus{}of}~\PY{o}{(}\PY{n}{canon}~\PY{n}{lambda}\PY{o}{)}~\PY{n}{lambda}\PY{o}{.}
    \end{input}
  \end{sentence}
\end{alectryon}