### 2022-07-02 doc/JFLA2022/Slides/HprimePhi0OmegaClosed.tex

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      \PY{k+kn}{Lemma}~\PY{n+nf}{H\PYZsq{}\PYZus{}Phi0\PYZus{}omega\PYZus{}exact\PYZus{}formula}~\PY{n+nv}{k}~\PY{o}{:}\nl
      ~~\PY{n}{H\PYZsq{}\PYZus{}}~\PY{o}{(}\PY{n}{E0phi0}~\PY{n}{E0omega}\PY{o}{)}~\PY{n}{k}~\PY{o}{=}\nl
      ~~~~\PY{k+kr}{let}~\PY{n+nv}{F}~\PY{n+nv}{f}~\PY{n+nv}{i}~\PY{o}{:=}~\PY{n}{iterate}~\PY{n}{f}~\PY{o}{(}\PY{n}{S}~\PY{n}{i}\PY{o}{)}~\PY{n}{i}\nl
      ~~~~\PY{k+kr}{in}~\PY{k+kr}{let}~\PY{n+nv}{g}~\PY{n+nv}{k}~\PY{o}{:=}~\PY{n}{S}~\PY{o}{(}\PY{l+m+mi}{2}~\PY{o}{*}~\PY{n}{k}\PY{o}{)\PYZpc{}}\PY{n}{nat}\nl
      ~~~~~~~\PY{k+kr}{in}~\PY{n}{iterate}~\PY{n}{F}~\PY{n}{k}~\PY{n}{g}~\PY{n}{k}\PY{o}{.}~
    \end{input}
  \end{sentence}
\end{alectryon}