### 2022-09-04 doc/chapter-encoding.tex
sart a new chapter on Goedel encoding (after Russel)

\chapter{G\"{o}del's Encoding (in construction)}
\label{chap:encoding}