### 2023-02-16 doc/chap-intro-goedel.tex
small correction in latex document

\chapter{Introduction}

This part contains comments, example  and exercises about Russel O'Connor's work on G\"{o}del's first incompleteness theorem.  This work is maintained by Coq community~\cite{CoqCommunity} volunteers: (projects Goedel~\cite{Goedel} and Hydra-battles~\cite{HydraBattles}).
The main reference to this work is  Russel O'Connor's article~\cite{OConnor05}, which we strongly encourage the reader to 
consult regularly.

Some changes are made to the aforementionned libraries, mainly because of the recent evolution of coq and its libraries. Nevertheless, the definitions, lemmas and theorems of the original contribution have been kept in this new release.

\emph{\color{red!80}These maintenance and documentation 
jobs have just started, and will probably be long to complete. Help is welcome!}

\section{File contents}

All Russel O'Connor's files have been stored in two directories, in order to simplify packages maintenance.

\begin{itemize}
\item \texttt{theories/goedel/}: Proofs which depend on 
\texttt{CoqPrime} package.
\item  \texttt{theories/ordinals/Ackermann/} : all the rest:
 definition of primitive recursive functions, first-order  logic,
Peano Arithmetics, G\"{o}del's encoding.
\item Some additions we made: examples, exercices, new notations, etc.,  are stored in a specific directory \texttt{theories/ordinal/MoreAck/}.
\end{itemize}

\begin{todo}
Add information on recent developments on formal 
proofs of G\"odel  incompleteness theorems. Justify the decision of working on \emph{this} development.  
\end{todo}

\section{Warning}
Russel O'Connors contribution contains more than 42 KLoc.
Since its construction, \coq{},  its libraries and recommended style have evolved a lot. We have just started to ``modernize'' this code. We apologize for provisional inconsistencies of presentation (code and documentation).