### 2023-07-03 doc/chapter-fol.tex
improve chapter-fol.tex

\chapter{First Order Logic (in construction)}
\label{chap:fol}

\section{Introduction}

This chapter is devoted to the presentation of  data structures for representing terms and first order formulas over a ranked alphabet, and the basic functions and predicates over these types, more precisely:

\begin{itemize}
\item Abstract syntax of terms and formulas over a ranked alphabet composed of function and relation symbols.
\item Induction principles over terms and formulas.
\item Definition and main properties of substitution of terms to variables.
\end{itemize}

Although all the following constructions come directly from Russel O'Connor's work~\cite{Goedel}, we introduced minor (mainly syntactic) changes  to
take into account recent changes in \coq(new constructions, tactics, notations, etc.).

\section{Data types}

\subsection{Languages}

A \emph{language} is a structure composed of relation and function symbols, each symbol is given an \emph{arity} (number of arguments)\,\footnote{As suggested by Russel O'Connor in~\cite{OConnor05}, we consider two arity functions instead of a single function defined on the sum type \texttt{Relations + Functions}.}.

From~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol}
\inputsnippets{fol/LanguageDef}

\subsubsection{Example: $L$, a toy language}
In order to show a few simple examples of statements and proofs, we define a small language with very few symbols:
two constant symbols: $a$ and $b$, three function symbols $f$, $g$ and $h$  (of respective arity 1, 1 and 2), three propositional symbols $A$, $B$ and $C$, two one-place predicates symbols $P$ and $Q$, and a binary relational symbol $R$.

From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}.

\inputsnippets{FolExamples/ToyDef}

\begin{remark}
  \label{rem:underscores}
  The constructors of types \texttt{Rel} and \texttt{Fun} are suffixed by an underscore, in order to reserve the names \texttt{a}, \texttt{f}, \texttt{h}, \texttt{R}, etc. to the functions which build terms and formulas (please look at Sect~\ref{sec:fol-term-notations} and \ref{sec:fol-atomic-notations}).
\end{remark}

\subsection{Terms}

Given a language $L$, we define the type of \emph{terms} and
$n$-\emph{tuples of terms} over $L$.

\inputsnippets{fol/TermDef}

\begin{remark}
This representation of terms uses mutually inductive data-types instead of lists or vectors of terms. Please see also Remark~\vref{hydra:mutually-inductive-vs-lists}.
\end{remark}

\begin{remark}[Variables]
In O'Connor's formalization of first-order logic,  variables are 
just natural numbers, and the conversion from
\texttt{nat} to \texttt{Term L} is the constructor (\texttt{@var $L$}).
Although other choices may be considered : PHOAS, de Bruijn indices, etc,  we still the data structures of~\cite{Goedel}, in order not to break long proof scripts which use this representation (please look at Section 2 of~\cite{OConnor05} for a related discussion).

\end{remark}

\subsubsection{Examples}
\label{sect:folTermExamples}
\label{sec:fol-term-notations}

Let us build a few \gallina terms over our toy language, 
respectively corresponding to the terms $a$,
$f(a)$, $h(f(a),a)$, and $h(f(v_0),g(v_1))$.

First, in order to make terms on $L$ more readable, we introduce a few abbreviations.

From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}.

\inputsnippets{FolExamples/termAbbreviations}

\inputsnippets{FolExamples/TermExamples1}

The following ``term'' \texttt{t4} is not well formed, since the arity of $h$ is not respected\,\footnote{Strictly speaking, it's not a (well typed) term!}.

\inputsnippets{FolExamples/TermExamplesFail}

\subsubsection{Other Languages}

\begin{todo}
Link to the chapter which presents \texttt{LNT} and \texttt{LNN}.
\end{todo}

\subsection{First-order formulas}

The type of first order formulas over $L$ is defined 
in~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} as an inductive data type, with a limited set of basic constructions:
\emph{term equalities} $t_1=t_2$,
\emph{atomic propositions} $R\;t_1\;\dots\;t_n$, where $R$ is a relation symbol of arity $n$,
\emph{implications} $A \rightarrow B$,
\emph{negations} $\sim\,A$,
and \emph{universal quantifications} $\forall\,i,\;A$.

\vspace{6pt}

\noindent From~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol}

\inputsnippets{fol/FormulaDef}

\begin{remark}

In~\cite{Goedel}, no \emph{constructors} of type \texttt{(Formula L)} are associated with
disjunction, conjonction, logical equivalence and existential quantifier. These constructs are formalized through \emph{definitions} in terms of \texttt{impH}, \texttt{notH} and \texttt{forallH}\,\footnote{Please keep in mind that we are considering a classical logic.}.

\inputsnippets{fol/FolFull, fol/folPlus}

This convention allows the user to reduce to 5 (instead of 10) the number of cases in ''\texttt{match F with \dots}'' terms. On the other hand, some computation may expand a connective like $\vee$ or
$\wedge$, or an existential quantification into a ``basic'' formula (see Sect.\vref{sect:fol-issue}).
\end{remark}

\subsection{Examples}

Let us give a few examples of first-order formulas over $L$.
\label{fol:examplesf1f2f3}
\begin{description}
\item[F1] $R\;a\;b$
\item[F2] $\forall v_0\; v_1, R\;v_0\;v_1 \arrow R\;v_1\;v_0$
\item[F3] $ \forall v_0, v_0=a \vee \exists\;v_1,\, v_0= f(v_1)$
\item[F4] $(\forall v_1, v_0 = v_1) \vee \exists\,v_0\,v_1, v_0 \not= v_1$
\item[F5] $v_0 =a \vee v_0 = f(v_1)$
 \item[F6] $\forall v_0,\,\exists v_1, v_0= f(v_1)\wedge v_0\not= v_1$
\end{description}

Let us now define these formulas as terms of type
\texttt{(Formula L)}.

First, we define abbreviations for atomic formulas over $L$.

\noindent From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}

\inputsnippets{FolExamples/toyNotationForm}

\label{sec:fol-atomic-notations}
\label{sect:folFormExamples}

\inputsnippets{FolExamples/FormExamples}

% The following interactions show how the formulas 
% of~\ref{sect:folFormExamples} may be 
% parsed or printed.

% \noindent From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}

% \vspace{6pt}

% \inputsnippets{FolExamples/toyNotationForm2}

\subsubsection{Bound variables}

In~\cite{Goedel}, there is no De Bruijn encoding of bound variables (see also~\cite{OConnor05}).

For instance, the term \texttt{(var 0)}
occurs both freely and  inside the scope of a quantifier in the
formula \texttt{F4} \vpageref[above]{sect:folFormExamples}.

The following example shows two formulas which share the same structure, are logically equivalent, but are not Leibniz equal.

\vspace{4pt}

\noindent From~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}

\inputsnippets{FolExamples/boundVars}

  \begin{todo}
   Link to the lemmas which attest the equivalence of these formulas (properties of substitution, logical equivalence).
  \end{todo}

  \index{ackermann}{Projects}
  \begin{project}
 Define a [P]HOAS representation for FOL terms and formulas.
Could we avoid  to  break some proof scripts?   
  \end{project}

\section{A notation scope for first-order terms and formulas}
\label{sect:fol-notations}

 We use \coq's \texttt{Notation} features to print and parse terms and formulas  in a more readable form.
To this purpose, we build \texttt{fol\_scope}, a  notation scope
where the main connectives and quantifiers get a syntax close to \coq's.  Additionnally, a term of the form \texttt{(@var \_ $i$)} is
just printed and parsed  \texttt{v\#$i$}.

\inputsnippets{fol/folScope1}

The \texttt{\%fol} delimiter 
allows the user to distinguish \texttt{FOL} connectives from their \coq equivalent.
\emph{We discourage the reader from \emph{opening} \texttt{fol\_scope} and similar scopes : \texttt{nn\_scope}, \texttt{nt\_scope}, which would make disappear the \texttt{\%fol} suffix from the first-order formulas}.

\vspace{6pt}

\noindent From \href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol}

\inputsnippets{FolExamples/toyNotationForm2}

\subsection{The issue with derived constructions}
\label{sect:fol-issue}

The connectives and quantifiers $\vee$, $\wedge$, $\exists$, etc. may raise an issue when printing computed formulas.
For instance, a formula like $F \wedge G$ could be transformed into $\sim(\sim F \vee \sim G)$,  and even
into $\sim(\sim\sim F \arrow \sim B)$, which would 
cause serious problems of readability.

In such a case, we propose to print such a formula as $F \wedge' G$, to make it syntactically distinct but very similar to $F \wedge G$.

\inputsnippets{fol/folScope2}

The following examples show how the primed connectors 
and quantifiers behave with respect to convertibility and 
input/output.

 \inputsnippets{FolExamples/toyNotationForm3}

\inputsnippets{FolExamples/toyNotationForm4}

\begin{remark}
In some situations (like in the proof of \texttt{PrfEx4}~\vref{prfex4})
the user may be puzzled when a formula [s]he typed explicitely
\texttt{$\sim A$ ->$\sim B$}  will be printed 
\texttt{$A$  \textbackslash/' $\sim B$}. We will try to fix this issue.
\end{remark}

\section{Computing and reasoning on first-order formulas}

\subsection{Structural recursion on formulas}

Structural induction/recursion principles are generated from
\texttt{Term}, \texttt{Terms} and \texttt{Formula}'s definition, for instance:

\inputsnippets{fol/TermSchemes, FolExamples/FormulaRect}

\subsubsection{Free variables}
The functions \texttt{freeVarT} [resp. \texttt{freeVarTs},  and
\texttt{freeVarF}] compute the multiset (as a list with possible repetitions) of the free occurrences of variables in a term [resp. a vector of terms, a formula].

\inputsnippets{folProp/freeVarT}

Concerning formulas, the treatement of binding is realized 
whith the help of the \texttt{List} library function \texttt{remove}.
\inputsnippets{folProp/freeVarF, FolExamples/freeVarEx}

\begin{remark}
  Note that \texttt{freeVarF} is defined by cases over the \emph{basic} connectives. Formulas with contain \texttt{iffH} or
\texttt{ifThenElseH} are expanded before the application of
\texttt{freevarF}, and the 
list returned by \texttt{freeVarF} may contain redundancies.

If we want to get the \emph{set} of variables wich occur freely 
in a formula $F$, we may use the function \texttt{List.nodup}.

\inputsnippets{FolExamples/freeVarDup}

\end{remark}

\subsubsection{Closing a formula}
Function \texttt{freeVarF} is used in 
the  function \texttt{close}, which  universally quantifies  all the free variables of a formula.

\noindent From \href{../theories/html/hydras.Ackermann.folProp.html}{Ackermann.folProp}

\inputsnippets{folProp/closeDef}

\noindent From \href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}

\inputsnippets{FolExamples/closeExample}

\begin{remark}
  The function \texttt{close} applies \texttt{freeVarF} and
  \texttt{List.nodup}  in order to add a sequence of universal
  quantifications (\texttt{allH $i_1\, \dots\, i_k$}), in an order determined by the actual implementation of these functions. It may be interesting to check whether the proof of properties of \texttt{close} depend or not from this implementation.
\end{remark}
% \subsubsection{Computing a fresh variable}

% It is easy to get a new variable, which

\subsection{Decidability of equality}
% Deciding whether two formulas are equal is a part of the verification of a proof (for instance in the implementation of the \emph{assumption} rule).

Let $L$ be a language, and let us assume that equality 
of function and relation symbols of $L$ are decidable.
Under this assumption, equality of terms and formulas over $L$ is decidable too.

Because of dependent types, the proofs are quite long and technical. The reader may consult them in \href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol}

\inputsnippets{fol/formDec1, fol/formDec2,
fol/formDec3,fol/formDec4,fol/formDec5}

\begin{remark}
Please note that \texttt{term\_dec}, \texttt{terms\_dec} and
\texttt{formula\_dec} are \emph{opaque}.

The function \texttt{formula\_dec} is mainly used in
\href{../theories/html/hydras.Ackermann.PA.html}{Ackermann.PA}, in order to check whether a given formula belongs to the axioms of Peano arithmetic.

\begin{todo}
  Look for the use of \texttt{open}  (in codePA)
\end{todo}

\end{remark}

\subsection{Variables and substitutions}

The substitution of a term to the free occurrences of a given variable $v$ is at the heart of the implementation of universal
quantifier elimination.

Since free and bound occurrences of a variable $i$ are represented the same way, much care should be taken in programming the substitution of a term to a variable in order to avoid \emph{variable capture}.

Substitution of a term $t$ to all the occurrences of a variable $x$ in a term or a vector of terms is easy to define as a pair of mutually structurally recursive functions.

\inputsnippets{folProp/subsTdef}
\inputsnippets{FolExamples/substTExample}

Concerning formulas, it could be tempting to define substitution the same way.
\inputsnippets{BadSubst/BadSubstFdef}

Let us consider for instance the formula $F = \forall\,v_1, \,\exists\,v_2,\,v_1\not=f(v_2)$ (satisfiable if we take for instance $f$ to be interpreted as the successor function on natural numbers).

If we eliminate the universal quantifier by substituting in
the sub-formula $F_1=  \exists\,v_2,\,v_1\not=f(v_2)$ the free
occurrences of $v_1$ with $f(v_2)$, our naive implementation of
\texttt{substF} returns the absurd proposition $\exists\,v_2,\,f(v_2)\not=f(v_2)$. We  say that the free occurrence of $v_2$ in the term $f(v_2)$ has been \emph{captured} by the binding $\exists\,v_2,\dots$.

\inputsnippets{BadSubst/BadSubstFexample}

In this example, we could obtain a correct result, if
\begin{enumerate}
\item We consider a \emph{fresh} variable, \emph{i.e.}  different from $v_1$ and $v_2$, say for instance $v_3$,
\item we substitute $v_3$ to $v_2$ in $F_1$ which results in
   $F_2= \exists\,v_3,\,v_1\not=f(v_3)$
\item we substitute the term $f(v_2)$ to $v_1$ in $F_2$, which gives us 
  $\exists\,v_3,\,f(v_2)\not=f(v_3)$.
\end{enumerate}

The notion of fresh variable is implemented through a function
\texttt{newvar (l: list nat) : nat} which returns a number which doesn't belong to $l$.

But the following attempt fails, because the renaming of a  variable in a 
sub-formula of a formula $F$ is not structurally smaller than $F$.

\inputsnippets{BadSubst/BadSubstF2}

%ICI

%\inputsnippets{FolExamples/substExample1}

% Thus, the function (\texttt{fun f => substF L f i t}) doesn't commute with all \texttt{Formula}'s constructors.

% \inputsnippets{FolExamples/substExample2}

% From this little example, we can guess that the function
% \texttt{substF} which is assumed to compute the substitution in a formula $F$ the substitution of a term $t$ to the free occurrences of a variable $v$ will not be implemented as a direct structural recursive function. For instance, the formula $G$, obtained through a variable renaming, is not a strict subterm of $F$.

% Thus, we cannot expect do define simply this operation by 
% structural recursion over formulas. 
% In~\cite{Goedel}, the approach is to define a \emph{measure}
% which maps every formula to a natural number, then define substitution in such a way that the substitution of a term to a free variable in a strict sub-formula  of a formula $F$ is smaller
% (w.r.t. this measure) than $F$. 
% Then substitution could be defined by well-founded recursion.

Fortunately, \coq allows us to define functions by well-founded recursion, and in particular with the help of a \emph{measure} mapping every formula to an already known well-founded type.

% To fix this issue, we can associate  a measure (\emph{e.g.} a natural number) to each formula, such that the renaming of a strict subformula $F$ is strictly smaller that $F$.

\subsubsection{Depth of a formula}

The function \texttt{depth} computes the \emph{depth} of  any formula, \emph{i.e} the height of the sub-tree made
by erasing all nodes but those nodes labelled with
\texttt{allH}, \texttt{impH} and \texttt{notH}.

\inputsnippets{fol/depthDef}

\begin{remark}
  The depth of a formula takes into account its abstract syntax tree \emph{with respect to the base connective and quantifiers : $\arrow$, $\sim$ and $\forall$}.
    Formulas which contain $\vee$, $\wedge$, $\exists$, etc. are translated into
    basic formulas before the computation of their depth. In the example below, 
the conjunction is translated into a bigger term than the disjunction.

\inputsnippets{FolExamples/DepthCompute}
\end{remark}

\subsubsection{Induction on depth}

Lemma \texttt{fol.Formula\_depth\_rec} is the basic
induction principle based on depth.

\inputsnippets{FolExamples/depthRec}

% \begin{todo}
% Motivate the induction principles based on depth. Compatible with term substitution and universal quantifier elimination.
% \end{todo}

\inputsnippets{FolExamples/depthRecDemo}

\begin{todo}
 Look for the principles which are really used in Ackermann or/and Goedel libraries, and comment them.
 Maybe skip the helpers (unused in other files)
\end{todo}

The library~\href{../theories/html/hydras.Ackermann.fol.html}{Ackermann.fol} contains several derived induction principles, applied 
throughout \texttt{Ackermann} and \texttt{Goedel} projects.

Let us for instance have a look  at \texttt{Formula\_depth\_ind2} which helps  to prove a goal $(P\;a)$ by generating five sub-goals.

\inputsnippets{FolExamples/depthRecDemo2}

\begin{itemize}
\item Goals $1$ to $4$ correspond to  usual proofs by structural induction (without refering to \texttt{depth}).
\item Goal $5$ is associated with a universal quantification $f=\forall\,v,a$. In this case, we have to prove that $P\;b$ holds for any formula $b$ which has a depth strictly less than $f$. Such a $b$ may for instance be the result of replacing the free occurrences of $v$ in $a$ with any term $t$.

  \begin{todo}
   Make a link to an appropriate example.
  \end{todo}
\end{itemize}

\subsection{A correct definition of substF}

Substitution of a term to (free- occurrences of a variable 
in a formula is defined in a section of \href{../theories/html/hydras.Ackermann.folProp.html}{Ackermann.folProp}.
The definition itself takes 200 lines of \coq code, so we will only comment its structure.

Fortunately, the reader may skip a few complex definitions 
(often because of dependent pattern matching), whose purpose
is twofold:
\begin{itemize}
\item Ensure that the substitution in a formula $F$ of a term $t$ to a variable $x$ returns a formula of the same depth as $F$,
  which allows to define a function by well-founded recursion on depth.
  \inputsnippets{folProp/substFHelp, folProp/substFDef}
\item Prove a few \emph{equations} which will be used in further proofs.

\inputsnippets{folProp/subFormulaEqual}
\inputsnippets{folProp/subFormulaRelation}
\inputsnippets{folProp/subFormulaImp}
\inputsnippets{folProp/subFormulaNot}
\inputsnippets{folProp/subFormulaForall}

\item Similar equations are also proved for derived connectors and quantifiers, for instance:

\inputsnippets{folProp/subFormulaAnd}
\inputsnippets{folProp/subFormulaExist}

\end{itemize}

Let us look at a few examples. Despite the complexity of its definitoin, the function \texttt{substF} behaves well with respect with computations.

\inputsnippets{FolExamples/substExample1}

\begin{project}
  Is it possible to get a more readable definition of \texttt{substF} using the \emph{Equations} plug-in~\cite{sozeau:hal-01671777} ? 
\end{project}

\subsection{Multiple substitutions}

The function \texttt{subFormula} defined in
\href{../theories/html/hydras.Ackermann.subAll.html}{Ackermann.subAll} allows to substitute a term $t_i$ to ecah free occurrence of the variable $v_i$ in a formula $F$. The dependance of $t-i$ from $v_i$ is given through a function from \texttt{nat} to
(\texttt{Term L}).

\inputsnippets{FolExamples/subAllExample1}

\section{Proofs}

\subsection{Proof trees}

Proof trees in first-order logic are the inhabitants of the \texttt{Prf $L$} 
inductive type, defined in \href{../theories/html/hydras.Ackermann.folProof.html}{Ackermann.folProof} and displayed in Figure~\vref{fig:prf-def}.

Please note that the following constructions are parametrized with an
arbitrary language $L$, declared in the \texttt{ProofH} section.

\inputsnippets{folProof/prelude}

\begin{figure}[p]
  \centering
\inputsnippets{folProof/PrfDef}  
  \caption{Definition of the type \texttt{Prf} of proof trees}
  \label{fig:prf-def}
\end{figure}

\subsubsection{\texttt{Prf}'s type}

The type  \texttt{Prf $l$ $F$}  is the type of ``proof trees of the formula $F$, where $l$ is the list of assumptions used in the proof, enumerated left-to-right (\emph{i.e.} the fringe of the proof tree)''.

In the rest of this section, we comment  every one of 
\texttt{Prf}'s 14 constructors,
 and give simple examples of their application.

\subsubsection{Warning}

We won't respect the order in which \texttt{Prf}'s constructors
of type \texttt{folProof}  
are enumerated in \texttt{folProof.v} (see Figure~\ref{fig:prf-def}). 
Instead, we preferred to present 
these constructors in an order inspired by a sequence
 of simple examples. On the other hand, we didn't change this order in 
\texttt{folProof.v}, in order not to break complex proofs by pattern-matching.

For more information on Hilbert proof system, you may consult
\url{https://en.wikipedia.org/wiki/List_of_Hilbert_systems}.

\subsubsection{The axiom rule: \texttt{AXM}}

Let $A$ be a formula on $L$.  The \texttt{AXM} rule builds a proof-tree
of $A$ which uses \emph{exactly} the singleton list \texttt{[$\,A\,$]}.

\noindent From \href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}

\inputsnippets{FolExamples/PrfEx1}

\subsubsection{Modus Ponens: \texttt{MP}}

Let $A$ and $B$ be two formulas on $L$, $\textit{Axm1}$
 and $\textit{Axm2}$ two sequences of formulas.
If we have 
two proof trees of respective types \texttt{Prf Axm1 $A\arrow B$} and  \texttt{Prf Axm2 $B$}, then we build a proof tree
for $B$ whose fringe is the concatenation of $s_{AB}$ and $s_A$.

The following proof script is a quite naive application of 
\texttt{AXM} and \texttt{MP}.

\inputsnippets{FolExamples/PrfEx2}

We can make this proof script shorter, using existential variables.

\inputsnippets{FolExamples/cutMP}

\inputsnippets{FolExamples/cutMP2}

\subsubsection{Hilbert's axioms for implication: \texttt{IMP1} and \texttt{IMP2}}

\inputsnippets{FolExamples/PrfEx3}

\begin{remark}
\label{sec:remark-on-hilbert-systems}
One may think that this proof 
is quite clumsy. \emph{Right}.
But we must recall that \texttt{Prf} is a basic Hilbert-like
proof system, which will make it easier to study.
In the next chapter, we will consider a derived proof system 
where the \emph{deduction theorem} will allow us to consider shorter and more natural proofs.
\end{remark}

\begin{exercise}[**]
  Is it possible to build a term of type
\texttt{Prf L [A] B -> Prf L [] (A->B)\%fol} ?
\end{exercise}

\subsubsection{The rule of contraposition: \texttt{CP}}

The only rule about the \texttt{notH} connective is
the \emph{contraposition rule}.

The following script shows that \texttt{CP} entails the 
derived rule of \emph{proof by contradiction}.

\label{prfex4}

\inputsnippets{FolExamples/PrfEx4}

\begin{remark}
Same remark as~\ref{sec:remark-on-hilbert-systems}.
\end{remark}

\begin{exercise}
  Replay this proof with pen and paper!
\end{exercise}

\subsubsection{Rules about the universal quantifier: \texttt{FA1},
\texttt{FA2} and\texttt{FA3}}

\inputsnippets{FolExamples/PrfEx5}

\inputsnippets{FolExamples/PrfEx6}

\inputsnippets{FolExamples/PrfContrex7}

\inputsnippets{FolExamples/PrfEx8}

\subsubsection{Axioms for Equality: \texttt{EQ1} to \texttt{EQ3}}

The following proof applies \texttt{FA1} in order to build a proof
of \texttt{Prf L [] (t=t)\%fol}  for any $t$\,\footnote{Please note that the quantification on $f$ is \coq's. Thus, \texttt{eq\_ref} is kind of scheme.}

\inputsnippets{FolExamples/eqRefl}

\subsubsection{Axioms schemes for Equality: \texttt{EQ4} and \texttt{EQ5}}

The constructors \texttt{EQ4}  and \texttt{EQ5} build an infinite number of axioms, parameterized by a  relation  or function symbol.

The function \texttt{AxmEq4} generates a formula with $n$ pairs of free variables (with $n$ the arity of the considered relation symbol).

\inputsnippets{FolExamples/PrfEx910}

Please note that \texttt{EQ4} uses a \emph{sequence} of variables generated by \texttt{AxmEq4}. Any other sequence may cause 
\texttt{EQ4} to fail.

\inputsnippets{FolExamples/PrfContrex9}

The following script shows that \texttt{EQ5} is to function symbols 
what \texttt{EQ4} is to relation symbols.

\inputsnippets{FolExamples/PrfEx11}

\section{Concluding remarks}

The type \texttt{Prf} is composed of very simple rules.
Nevertheless, the examples presented in the previous section seem to show that proving even simple theorems is not trivial at all.

Indeed, in the next chapter, we will consider a proof system
\texttt{SysPrf}, based on \texttt{Prf}, the properties of which will 
allow us to prove theorems in a very simpler way.

%%%%%%%%%%%%%%% New chapter!