### 2023-07-30 doc/chapter-natural-deduction.tex
corrections chap 13

\chapter{Natural Deduction (in construction)}

%%%%%%%%%%%%%% New chapter 
\section{Contexts as sets}

Let us look again at the proof scripts shown in~\vref{proof:PrfEx2} and \vref{proof:PrfEx3}.
\begin{itemize}
\item The statement of \texttt{PrfEx2} contains a sequence
of hypotheses with two occurrences of \texttt{A}. Moreover, the order in which the 4 hypotheses are listed is determined by the type of the constructor \texttt{MP}  (please look at
Figure~\vref{fig:prf-def}). It would be better to replace this
precise list of hypotheses with ``any list whose elements
belong to the set $\{A, A \arrow B; A \arrow B \arrow C\}$''.

\item In the proof of \texttt{PrfEx3} the \coq user would certainly ask ``How do we apply implication's introduction rule?''.
\end{itemize}

The answer to both questions in~\cite{Goedel}  is the definition of a proof system, derived from \texttt{Prf}, which considers
\emph{sets of hypotheses} (called \emph{systems} in ~\cite{Goedel}) instead of
\emph{list of hypotheses}, thus making abstraction of
the repetition and order of appearance of hypotheses in the context.

The new system is simply defined as below (for  a given language $L$)\,\footnote{In some shown snippets, arguments like $L$ may be or not be implicit (depending on the section they are extracted from). Please look at the \coq source.}.

\inputsnippets{folProof/SysPrfDef}

\subsubsection{Notations}

In the text, we may use the abbreviation $\Gamma \sdash A$ for
(\texttt{SysPrf $L$ $\Gamma$ $A$}) and
$\sdash A$ for
(\texttt{SysPrf $L$ Empty\_set $A$})

We may also omit obvious braces in some set expressions:

\begin{itemize}

\item $\Gamma, \Delta$ for $\Gamma\cup\Delta$,
\item $A, B, C$ for $\{A, B, C\}$,
\item $\Gamma, A,\dots B$ for $\Gamma\cup\{A \dots B\}$
\item \emph{etc.}
\end{itemize}

\begin{remark}
  The type \texttt{SysPrf} has sort \texttt{Prop}, which prevent us from
extracting the underlying proof tree and its fringe from a proof of $X\sdash f$. We only know that such a proof exists, but cannot get it automatically through a \coq function.
\end{remark}

\index{ackermann}{Projects}

\begin{project}
It would be nice (\emph{e.g.} for a better understanding of the proof of the deduction theorem) to be able to compute the  proof-tree built
by the proof of the deduction lemma. 
On a fresh branch of the project, please change the definition of
\texttt{SysPrf} and fix the errors this change could cause in the rest of the files. 
\emph{If ensuring compatibility with all the \texttt{Goedel} project is too long and/or difficult, you may just make the changes in separate modules with an ``informative'' \texttt{SysPrf} and limit the compatibility study  to the contents of basic modules like
\texttt{Deduction}, \texttt{folLogic}, etc.
Perhaps you will have to consider other implementation of finite sets of formulas (\emph{e.g.} lists).
}

\end{project}

\subsubsection{Examples}

In the following script, we use \texttt{PrfEx2}
as a witness for proving a set-based version of the 
original proof term, namely $A; A \arrow B; A \arrow B \arrow C \sdash C$.

\inputsnippets{FolExamples/SysPrfEx2}

\begin{exercise}

Prove the following lemma (without the handy lemmas from
\href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic} and their corollaries!).

\inputsnippets{FolExamples/MPSys}

\end{exercise}

\subsection{Using properties of sets}

The following three lemmas, from 
\href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic} are direct consequences of \texttt{SysPrf}'s definition.

\inputsnippets{folLogic/Axm}

\inputsnippets{folLogic/SysExtend}

The \emph{rule of implication elimination} is derived from \texttt{Prf}'s 
\emph{modus ponens} \texttt{MP}.  Since all elements of the fringe 
\texttt{x} [resp. \texttt{x1}] of the proof tree \texttt{px} [resp. \texttt{px1}] belong to T, so are the elements of the fringe \texttt{x++x1} of
(\texttt{MP \dots{}  px px1}).

\inputsnippets{folLogic/impE}

\section{The Deduction theorem}

The deduction theorem (proved in \href{../theories/html/hydras.Ackermann.Deduction.html}{Ackermann.Deduction} is a handy 
tool for proving a proposition $f\arrow g$ by pushing 
the hypothesis $f$ into the context (it corresponds roughly to 
the implication introduction rule in \coq).

\inputsnippets{Deduction/deductionThm}

\subsection{Sketch of proof}

We advise the reader to replay this proof on h.er.is computer
 in order to better understand its structure, which we will only comment briefly.

Let us assume the hypothesis \texttt{H: $T, g \sdash f$},
meaning that there exists some list $F$ whose elements 
belong to $T \cup \{g\}$, and a proof-tree $t$ of type
$F \vdash h$.

The heart of the proof is an induction on $t$ proving 
$F \cap T \sdash (g\arrow f)$ \footnote{Please forgive the implicit coercion from lists to sets!}.

Please note that the case $S=\texttt{nil}$ is common to 
many constructors of \texttt{Prf}, thus the proof script starts
with a study of this particular case, simply applied 11 times in the rest of the proof.

\section{Derived rules and natural deduction}

 The library 
 \href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic} contains many derived rules which allow the user to build proofs in a natural deduction style (with introduction and elimination rules).

%\subsection{Applications of the Deduction Theorem}

The \emph{rule of implication introduction} is a trivial application of the deduction theorem.

\inputsnippets{folLogic/impI}

The following lemma (corresponding to the \texttt{CP} constructor of type \texttt{Prf}) is proven with the help of \texttt{impE}, \texttt{impI}, 
\texttt{sysWeaken} and \texttt{CP}.

\inputsnippets{folLogic/contradiction}

We let the reader consult the proof of the following lemmas, or, much better, re-prove them as exercises.

\inputsnippets{folLogic/nnE}

\inputsnippets{ folLogic/nnI}

\inputsnippets{folLogic/cp1}

\inputsnippets{folLogic/cp2}

\subsection{Rules for derived connectives and quantifiers}

Let us keep in mind that the derived connectives: \texttt{orH}, \texttt{andH}, etc. and the existential quantifier \texttt{existH}  are defined
in terms of \texttt{impH}, \texttt{notH}, \texttt{forallH}.
By unfolding these definitions, we prove easily a few natural deduction rules for the derived symbols. 
For instance, the law of excluded middle for $f$: $ T\sdash \sim f \vee f$
for any $T$,  is just an abbreviation of $T \sdash \sim\sim{}f \arrow f$.

\inputsnippets{folLogic/noMiddle}

\begin{remark}
  The lines marked ``optional'' are just here in order to
temporarily deactivate the notation which print any formula 
of the form $\sim{}\!A \arrow B$ as \texttt{$A$ \disj' $B$}.
 These three lines can be safely removed.
\end{remark}

Let us go on \dots. We leave the completion of the proofs as an exercise. We encourage the reader to write h.er.is own proofs, then compare them with the scripts of \href{../theories/html/hydras.Ackermann.folLogic.html}{folLogic}.

\inputsnippets{folLogic/orI1}

\inputsnippets{folLogic/orI2}

\inputsnippets{folLogic/orE}

\inputsnippets{folLogic/orSys}

\inputsnippets{folLogic/andI}

\inputsnippets{folLogic/andE1, folLogic/andE2}

\inputsnippets{FolExamples/SearchSysPrf}

\subsection{Example: proof of Peirce's law}

For instance, let us prove Peirce's rule, \emph{i.e.}
$\sdash ((A \arrow B) \arrow A) \arrow A$ for any formulas $A$ and $B$.

 The proof in \coq is available in  ~\href{../theories/html/hydras.MoreAck.FolExamples.html}{MoreAck.FolExamples}.

\inputsnippets{FolExamples/PeirceProof}

Let us start with  an implication introduction. The judgement to prove becomes  $(A \arrow B) \arrow A \sdash A$.

\inputsnippets{FolExamples/step1}

Now, we may eliminate the disjunction 
in the instance $\sim A \vee A$ of the law of excluded middle.
The only non-trivial case is about $\sim A$.

\vspace{4pt}

Next goal is $(A \arrow B) \arrow A \sdash \sim\!A \arrow A$.
Please keep in mind  that our current notation system interprets
$\sim\!A \arrow A$ as a disjunction!

\inputsnippets{FolExamples/step2}

The rest of the proof is composed of basic proof steps.

\inputsnippets{FolExamples/step3, FolExamples/step4}

\index{ackermann}{Exercises}
\begin{exercise}
Prove, using the rules described in 
 \href{../theories/html/hydras.Ackermann.folLogic.html}{Ackermann.folLogic}, the famous \emph{drinkers theorem}:

$$\exists\,x,\; (D(x)\Longrightarrow \forall\,y,\; D(y))$$
\end{exercise}

where $D$ (for ``drinks'') is some predicate symbol of arity $1$.